% ===== SECTION 1: TRANSMITTAL LETTER =====
\section{\texorpdfstring{\color{MidnightBlue}Letter of Transmittal}{Letter of Transmittal}}

\begin{flushright}
\today
\end{flushright}

\vspace{0.5cm}

\noindent
\textbf{Rhode Island Department of Elementary and Secondary Education} \\
Division of Purchases \\
Statewide Transportation Services \\
Providence, Rhode Island

\vspace{0.5cm}

\noindent
\textbf{RE: RFI\#26006369 --- Statewide Student Transportation Services}

\vspace{0.5cm}

\noindent
\textbf{Dear Members of the Review Committee,}

\vspace{0.3cm}

Fejost LLC, doing business as \textbf{Sentry Management Solutions} (``Sentry''), is pleased to submit this response to RIDE's Request for Information regarding Statewide Student Transportation Services. We appreciate the opportunity to contribute to the development of an RFP that will shape the future of student transportation in Rhode Island for the 2027--2028 school year and beyond.

\vspace{0.3cm}

\subsubsection*{Who We Are}

Sentry Management Solutions is a New York--based transportation management and technology company founded in 2016. As a NYC and NYS Certified Minority Business Enterprise (MBE), Sentry specializes in technology-driven transportation management---delivering AI-powered routing optimization, real-time GPS dispatch, fleet analytics, and video monitoring solutions that give agencies centralized visibility and control across multi-vendor transportation networks. Today, Sentry serves as the prime broker for the MTA's Access-A-Ride paratransit program under an approximately \$181M contract, dispatching over 10,000 daily trips, tracking more than 15,000 vehicles, and managing roughly one million rides per year through our cloud-based platform---the largest MBE prime contract in MTA history.

\vspace{0.3cm}

\subsubsection*{Our Interest in the STS}

\begin{description}[style=nextline, leftmargin=1cm]
    \item[\textbf{Technology \& System Management:}] Sentry offers RIDE an enterprise-grade platform for routing optimization, real-time dispatch, fleet analytics, and video violation detection monitoring---capabilities that align directly with the 2025 legislative mandates and the Workgroup's efficiency goals.

    \item[\textbf{Vendor-Neutral Integration:}] Our technology is designed to operate across multiple bus operators and zones, providing RIDE with centralized visibility regardless of which contractors are awarded individual zones.

    \item[\textbf{Data-Driven Cost Reduction:}] Sentry's platform uses AI-powered route optimization and real-time ridership verification---including GPS geofencing, automated passenger counts, and dynamic routing algorithms---to eliminate deadhead miles, detect ghost riders through automated attendance reconciliation, and identify routes that can be rightsized from full-size buses to lower-cost vans, directly supporting RIDE's cost reduction and efficiency objectives.

    \item[\textbf{Video Violation Detection Readiness:}] Sentry has extensive experience deploying and managing onboard camera systems and telematics hardware---including Geotab GO9+ devices and Samsung tablet-based monitoring---across large fleets. Our platform is architected to ingest, process, and flag video violation events at scale, and we are prepared to deploy a fully compliant live digital video violation detection system well ahead of the July~1, 2027 mandate for new buses.
\end{description}

\vspace{0.3cm}

\subsubsection*{Authorized Representative}

The person authorized to represent Fejost LLC dba Sentry Management Solutions is:

\vspace{0.3cm}

\begin{tabular}{ll}
    \textbf{Name:}    & Max Zinner \\
    \textbf{Title:}   & Chief Strategy Officer \\
    \textbf{Phone:}   & 917-522-2323 \\
    \textbf{Email:}   & mzinner@sentryms.com \\
    \textbf{Address:} & 14 Bruckner Blvd, Unit 3B, Bronx, NY 10454 \\
\end{tabular}

\vspace{0.3cm}

\subsubsection*{Contact for Clarification}

\begin{tabular}{ll}
    \textbf{Name:}    & Max Zinner \\
    \textbf{Title:}   & Chief Strategy Officer \\
    \textbf{Phone:}   & 917-522-2323 \\
    \textbf{Email:}   & mzinner@sentryms.com \\
\end{tabular}

\vspace{0.3cm}

This response is submitted for informational purposes in accordance with the terms of RFI\#26006369. We understand that no award will be made as a result of this solicitation and that all information marked as proprietary shall be treated as confidential.

\vspace{1cm}

Respectfully submitted,

\vspace{1.2cm}

\rule{6cm}{0.4pt}\\
Max Zinner\\
Chief Strategy Officer\\
Fejost LLC dba Sentry Management Solutions\\
(844) 573-6879 $\cdot$ info@sentryms.com
