% ===== SECTION 2: COMPANY OVERVIEW =====
\section{\texorpdfstring{\color{MidnightBlue}Company Overview}{Company Overview}}

\subsection{Organization Summary}

\begin{tabularx}{\textwidth}{|>{\bfseries}l|X|}
\hline
\rowcolor{tableHeader}\multicolumn{2}{|l|}{\textcolor{white}{\textbf{Organization Summary}}} \\
\hline
Legal Entity Name          & Fejost LLC dba Sentry Management Solutions \\
\hline
Year Established           & 2016 \\
\hline
Type of Ownership          & Limited Liability Company (LLC) \\
\hline
Parent Company             & None --- Sentry is an independent, privately held company \\
\hline
Headquarters               & 14 Bruckner Blvd, Unit 3B, Bronx, NY 10454 \\
\hline
Total Full-Time Employees  & Approximately 50 \\
\hline
Primary Contact            & Max Zinner, Chief Strategy Officer \\
\hline
Phone / Email              & 917-522-2323 / mzinner@sentryms.com \\
\hline
Website                    & www.sentryms.com \\
\hline
\end{tabularx}

\vspace{0.5cm}

\subsection{Sentry's Role in Student Transportation}

Sentry Management Solutions is \textbf{not a bus operator}. We are a transportation technology and system management company that provides the platform, analytics, and operational oversight that state agencies and transit authorities need to manage large, multi-vendor transportation networks. Our role is analogous to the Statewide System Manager function RIDE envisions---we sit between the commissioning agency and the transportation providers, delivering centralized routing optimization, real-time dispatch and GPS tracking, fleet performance analytics, video monitoring, and compliance reporting across all vendors and zones from a single integrated platform.

This distinction is critical to RIDE's next-generation STS. The current model bundles system management with bus operations under a single contract. Sentry's experience demonstrates that separating technology and management from vehicle operations unlocks competition among bus operators on price and service quality while giving the state a vendor-neutral command center with full visibility into every route, vehicle, and rider. Our cloud-based platform already performs this exact function at a scale far exceeding Rhode Island's STS---managing over 10,000 daily trips and tracking more than 15,000 vehicles across a network of 150+ transportation providers for the MTA's Access-A-Ride paratransit program.

Founded in 2016 and headquartered in the Bronx, New York, Sentry is a NYC and NYS Certified Minority Business Enterprise led by CEO Steven Rivera and a core team of approximately 50 experienced transportation professionals. Our technology has been purpose-built for the complexities of public-sector transportation: multi-modal fleets, ADA compliance, IEP-driven routing constraints, dynamic address changes (McKinney-Vento), multilingual communications, and rigorous reporting requirements---all challenges that mirror RIDE's STS environment.

\begin{sidebar}{Core Capabilities}
\begin{description}[style=nextline, leftmargin=0.2cm]
    \item[\faIcon{route} Routing Optimization:] AI-powered routing engine that optimizes across multiple zones simultaneously, minimizing deadhead miles and ride times while respecting constraints such as IEP-mandated maximum ride durations, wheelchair accessibility, regional boundaries, and school bell schedules. Integrates with 12+ third-party dispatch, routing, and GIS systems.
    \item[\faIcon{headset} Real-Time Dispatch:] Cloud-based dispatch platform with live GPS tracking of all vehicles, automated ETA calculations, geofencing alerts for route deviations, and a multilingual call center operating at a 1:6.5 supervisor-to-agent ratio with 95\% same-day issue resolution. Currently dispatches 10,000+ trips daily with 95\% on-time pickup performance.
    \item[\faIcon{video} Video Violation Detection:] Proven deployment of onboard camera systems and telematics hardware (Geotab GO9+ devices, Samsung tablets) across large fleets. Platform architecture supports ingestion, AI-assisted analysis, and automated flagging of video violation events, with a complete processing pipeline from capture through evidence packaging for enforcement.
    \item[\faIcon{chart-line} Analytics \& Reporting:] Real-time vendor performance dashboards, KPI scorecards, ridership trend analysis, cost-per-route breakdowns, ghost rider detection through automated attendance reconciliation, and configurable compliance reporting. Platform maintains 99.999\% uptime on AWS GovCloud infrastructure (PostgreSQL Aurora, DynamoDB, Redis, S3) with FERPA and HIPAA compliance.
    \item[\faIcon{bus} Fleet Management:] Centralized tracking of vehicle inspections, preventive maintenance schedules, and lifecycle status across all providers in the network. Real-time monitoring of 15,000+ vehicles with automated alerts for maintenance milestones, safety compliance deadlines, and vehicle age/condition thresholds to support rightsizing decisions (bus vs.\ van assignment).
    \item[\faIcon{mobile-alt} Parent \& Driver Communication:] Multi-channel communication suite including a parent/rider mobile app with real-time GPS bus tracking, SMS and push notifications for ETAs and delays, a driver-facing tablet application, and web portals for school administrators. Supports communications in 9+ languages, with ridership verification and electronic signature capture at pickup and drop-off.
\end{description}
\end{sidebar}

\subsection{Relevant Experience}

Sentry's most directly relevant engagement is our role as the \textbf{prime broker for MTA Access-A-Ride}, New York City's ADA paratransit program. Under an approximately \$181M contract (August 2021--present), Sentry manages the technology platform and vendor coordination for the nation's largest paratransit system---dispatching over 10,000 daily trips, tracking more than 15,000 vehicles in real time, and coordinating a network of 150+ local transportation providers across all five NYC boroughs. This contract, generating \$70M+ in annual services, is the largest MBE prime contract in MTA history. The multi-vendor, multi-zone structure of Access-A-Ride is directly analogous to RIDE's STS: Sentry serves as the system manager providing centralized technology, routing, dispatch, and performance oversight while independent operators provide the vehicles and drivers.

Beyond MTA paratransit, Sentry has demonstrated versatility across multiple public-sector transportation programs. For the \textbf{NYC Department for the Aging (DFTA)}, we serve as the sole citywide provider of senior transportation services under a \$1.7M contract (January 2023--present), reducing wait times by more than 30\% through dynamic routing. For \textbf{VillageCare}, a Medicaid Managed Long-Term Care plan, Sentry has managed non-emergency medical transportation (NEMT) since 2018, handling thousands of daily trips with a 100\% transition of transport services to our platform. We also operated non-emergency patient transport for \textbf{NYC Health+Hospitals} (2019--2023) and managed specialized programs for the MTA including homeless individual transportation (\$600K, 1,000+ individuals in 4 months) and the overnight Warming Bus Program at subway stations (\$745K, 2023--present).

Across all engagements, Sentry has deployed GPS tracking hardware, onboard camera and telematics systems (Geotab GO9+ devices, Samsung tablets), and our proprietary cloud platform---maintaining a 99.999\% uptime record and a safety record of just 3.25 accidents per million miles driven. Our experience managing hardware deployment, video data ingestion, and compliance reporting across large fleets positions us to support RIDE's live digital video violation detection mandate from day one.

\begin{tabularx}{\textwidth}{|>{\bfseries}l|X|}
\hline
\rowcolor{tableHeader}\multicolumn{2}{|l|}{\textcolor{white}{\textbf{MTA Access-A-Ride --- Prime Broker for ADA Paratransit}}} \\
\hline
Client              & Metropolitan Transportation Authority (MTA), New York City \\
\hline
Contract Value      & Approximately \$181M (August 2021--present); \$70M+ annual services \\
\hline
Fleet / Scale       & 15,000+ vehicles tracked; 150+ transportation providers managed \\
\hline
Daily Volume        & 10,000+ dispatched trips per day; approximately 1 million rides per year \\
\hline
Technology Deployed & Cloud-based dispatch and routing platform (AWS GovCloud), real-time GPS tracking, AI-powered route optimization, vendor performance dashboards, Geotab GO9+ telematics, Samsung tablet driver interface, multilingual call center, mobile app with SMS notifications \\
\hline
Service Area        & All five boroughs of New York City \\
\hline
\end{tabularx}

\subsection{Understanding of Rhode Island's STS}

Sentry has carefully studied the current structure of RIDE's Statewide Transportation System, established by the General Assembly in FY~2009 under RIGL~\S~16-21.1 and phased in over three years (14~LEAs in 2010, 26 in 2011, all LEAs by 2012). The system today operates across seven zones serving approximately 3,812 students through 334 routes coordinated by the current Statewide System Manager, \textbf{Transpar}, on behalf of 49~LEAs plus out-of-state placements. RIDE's own \textit{Statewide Transportation System Overview} (October~2024) demonstrates that this system has grown by \textbf{34\%} in ridership since 2015---from 2,850 to 3,812 students---even as total Rhode Island student enrollment has \textit{declined} by 4\% (141,959 to 136,154). The system's FY~2025 cost exceeds \textbf{\$41~million}.

\vspace{0.3cm}

This demand growth is driven by three distinct student populations, each with unique routing and compliance requirements:

\begin{itemize}
    \item \textbf{PCCT (Private, Charter, Career \& Technical)} --- 50.87\% of riders. Regional boundaries under RIGL~\S~16-21.1-2 apply to this population. The Meeting Street example illustrates the consolidation advantage: statewide routing reduced buses serving that single school from 28 to 13---a 54\% reduction.
    \item \textbf{Special Education} --- 24.29\% of riders (234 of 334 routes). Required by federal law (IDEA); regional boundaries do \textit{not} apply. Routes must accommodate wheelchair accessibility, 1:1 aides, and IEP-mandated maximum ride times.
    \item \textbf{Homeless \& Foster Care (McKinney-Vento)} --- 24.84\% of riders. Required by federal law; regional boundaries do \textit{not} apply. Addresses change frequently, requiring dynamic route adjustment.
\end{itemize}

\vspace{0.3cm}

We are particularly attentive to the three bills passed into law in 2025 as a result of the Student Transportation Workgroup's recommendations:

\begin{itemize}
    \item \textbf{Increased Van Size} --- The expanded 10-passenger definition presents an opportunity to convert qualifying routes from higher-cost school buses to vans, reducing per-route costs while maintaining service quality. RIDE has already begun ``rightsizing'' by increasing vans and minibuses; 147 of 334 vehicles are already vans.
    \item \textbf{Temporary School Bus Driver Licenses} --- Abbreviated CDL training for border-state residents can accelerate recruitment pipelines, addressing the driver shortage that competes with Uber, DoorDash, RIPTA, FedEx, and Amazon for the same labor pool. Notably, RI's unique ``white card'' requirement---a multi-day course at CCRI plus DMV driving tests---remains a barrier for out-of-state CDL holders even with the new temporary license provision.
    \item \textbf{Live Digital Video Violation Detection Systems} --- The July 1, 2027 mandate for new buses (and July 1, 2032 for all buses) creates an immediate need for a technology partner capable of deploying, managing, and processing video violation data at scale.
\end{itemize}
