% ===== SECTION 3: BARRIERS TO ENTRY =====
\section{\texorpdfstring{\color{MidnightBlue}Inquiries Regarding Potential Barriers to Entry}{Barriers to Entry}}

\begin{sidebar}{RFI Reference}
Section C: Questions to Field --- Inquiries Regarding Potential Barriers to Entry (Questions 1--4)
\end{sidebar}

% ----- Question 1 -----
\subsection{Question 1}
\textit{``What, if anything, may prevent your company from bidding on this scope of work?''}

Sentry Management Solutions (Fejost LLC dba) is a technology and system management provider, not a bus operator. Whether we can bid on the eventual procurements depends entirely on how RIDE structures them. The following factors could prevent our participation:

\begin{itemize}
    \item \textbf{Fleet \& Drivers:} Sentry does not own or operate a bus fleet and does not directly employ CDL drivers. As a broker for MTA Access-A-Ride---coordinating 10,000+ daily trips (approximately 25\% of the program's 40,000+ daily volume)---our model is technology-enabled network management, not vehicle operation. Our participation depends on whether RIDE issues a separate technology/system management RFP or permits technology providers to partner with bus operators on the transportation RFP.
    \item \textbf{RI Presence:} Sentry does not currently maintain an operational footprint in Rhode Island. Our headquarters are in Bronx, New York, and our primary operations serve the New York City metropolitan area. If yard space, fuel access, or local infrastructure are preconditions for bidding, state-provided support (addressed in Question~4) would be critical to enabling our participation. However, our cloud-based platform (hosted on AWS GovCloud with 99.999\% uptime) can be deployed to any geography without physical infrastructure requirements for the technology layer itself.
    \item \textbf{Procurement Structure:} If the procurement is structured as a single RFP exclusively for vertically integrated bus operators---requiring a single vendor to provide buses, drivers, \textit{and} technology---Sentry would be unable to bid independently. Issuing separate RFPs for bus operations and for technology/system management would enable our participation and, critically, would increase the overall number of qualified bidders across all contract components. RIDE's current use of Transpar as a separate System Manager demonstrates that this separation is both feasible and operationally sound.
\end{itemize}

% ----- Question 2 -----
\subsection{Question 2}
\textit{``What contract structure(s) would make participation in RIDE's Statewide Transportation System (STS) more feasible for your organization? For example, would any of the following potential contract structures make it more likely that your organization would bid on this scope of work?''}

\vspace{0.2cm}

Sentry strongly encourages RIDE to consider the following contract structures, which would increase competition, attract specialized providers, and ultimately deliver better outcomes for Rhode Island students:

\begin{description}[style=nextline, leftmargin=0.5cm]
    \item[\faIcon{check-circle} \textbf{Separate Technology / System Management Contract:}] RIDE should procure the Statewide System Manager as a standalone technology contract, separate from bus operations. This contract would encompass routing optimization, real-time GPS dispatch, fleet analytics, video violation monitoring, automated billing, and parent communication. Separating technology from operations allows RIDE to select best-in-class technology independently of bus operator selection, ensures vendor-neutral oversight across all zones and all operators, and eliminates the inherent conflict of interest when operators self-report their own performance data. Sentry's experience managing our portion of the MTA Access-A-Ride program---where we serve as an independent technology broker coordinating 10,000+ daily trips within the nation's largest paratransit system (40,000+ daily trips total)---demonstrates exactly this model at a scale far exceeding the RIDE system.

    \item[\faIcon{check-circle} \textbf{Multiple Vendor Awards per Zone:}] Allowing multiple vendors per zone would enable technology providers like Sentry to partner with bus operators, bringing specialized routing and monitoring capabilities to zones where operators may lack in-house technology.

    \item[\faIcon{check-circle} \textbf{Allowing Subcontracting or Partnerships:}] Cross-zone partnerships would enable smaller or specialized firms to combine resources. Sentry could serve as the technology backbone for one or more bus operators across multiple zones, creating a more competitive field.
\end{description}

\begin{sidebar}{Recommendation to RIDE}
RIDE already operates with a separate System Manager---Transpar currently provides routing, program billing, vendor performance monitoring, and customer service support across all 49~LEAs. The forthcoming procurements are an opportunity to \textit{strengthen} this model by issuing a dedicated technology RFP for an upgraded platform with modern capabilities: data-driven routing optimization, real-time GPS dispatch, live video violation detection, and automated cost analytics. A next-generation System Manager would replace the current 700-check-per-year manual billing process, eliminate annual paper re-enrollment that contributes to the 550 ``ghost rider'' problem, and give RIDE independent performance data rather than relying on operator self-reports.
\end{sidebar}

% ----- Question 3 -----
\subsection{Question 3}
\textit{``What minimum route density or fleet size makes bidding on a zone more viable for your organization?''}

As a technology and system management provider, Sentry's viability is not constrained by route density or fleet size in the traditional sense. Our platform dispatches 10,000+ trips daily for MTA Access-A-Ride (approximately 25\% of the program's 40,000+ daily volume)---it scales from 29 routes (Zone~D) to 334 routes (statewide) with equal efficiency. We address this question from two perspectives:

\begin{itemize}
    \item \textbf{As technology provider:} There is no minimum route density or fleet size. Our cloud-based platform is zone-agnostic, scales linearly, and carries marginal cost near zero for each additional route. In fact, deploying the platform across \textit{all} zones simultaneously is most efficient, because cross-zone data enables system-wide optimization---identifying, for example, that a vehicle staged in Zone~B could serve a nearby Zone~C route with less deadhead than the Zone~C depot assignment.
    \item \textbf{If partnering with a bus operator:} The partner operator's fleet and driver minimums would govern. Industry benchmarks suggest that 25--30 routes represents a viable minimum for a dedicated depot to achieve economies of scale in driver supervision, maintenance, and fueling. Subdividing larger zones (e.g., splitting Zone~A1's 122 routes into two zones of 60+) would remain well above this threshold while lowering the capital barrier for new entrant operators.
\end{itemize}

% ----- Question 4 -----
\subsection{Question 4}
\textit{``What support (e.g., yard space, fuel access, maintenance facilities, etc.) could the State provide or assist with that would make it more likely for your organization to bid?''}

\begin{description}[style=nextline, leftmargin=0.5cm]
    \item[\faIcon{database} \textbf{Data Access:}] Providing prospective bidders with anonymized route data, FERPA-compliant student origin/destination data, current schedules, zone boundary GIS shapefiles, and historical performance metrics (on-time rates, deadhead times, road call frequencies) would allow technology providers to demonstrate measurable value during the procurement process. With this data, Sentry could present optimized routing scenarios showing projected deadhead reductions, vehicle count savings, and cost-per-student improvements---giving RIDE a concrete basis for evaluating technology platforms before contract award.

    \item[\faIcon{server} \textbf{Technology Infrastructure:}] State-provided or state-facilitated API access to existing RIDE systems---particularly the eRIDE student information system and IEP/special education databases---would accelerate deployment and reduce implementation risk. Sentry's platform is built on open REST APIs and has integrated with 12+ third-party dispatch, routing, and GIS systems. Pre-procurement API documentation and sandbox access would allow technology vendors to demonstrate working integrations rather than making promises.

    \item[\faIcon{warehouse} \textbf{Yard Space \& Facilities:}] While Sentry does not require yard space for its technology operations, state-owned or state-leased yard space made available to new entrant bus operators would significantly reduce the capital barrier for operators partnering with technology providers like Sentry. Incumbent operators have established depots and fueling infrastructure that took years to build; state-provided facilities would level the playing field and encourage new bidders who bring operational capability but lack Rhode Island real estate.

    \item[\faIcon{gas-pump} \textbf{Fuel Access:}] Centralized fueling contracts or state-negotiated bulk purchasing arrangements would reduce per-gallon costs for smaller operators and eliminate the competitive advantage that incumbents derive from established fueling infrastructure. This is particularly impactful given RIDE's 334-vehicle fleet consuming fuel across seven zones---a centralized contract could yield volume discounts that individual small operators cannot negotiate independently.

    \item[\faIcon{clock} \textbf{Extended Transition Period:}] The current timeline anticipates approximately 8 months from contract award (November 2026) to service commencement (July 1, 2027). For a technology deployment, Sentry can meet this timeline---our platform has been deployed to new transit networks in as few as 12 weeks. However, for bus operators who must recruit RI-licensed drivers, acquire or reposition vehicles, and establish depot infrastructure, 8 months is aggressive given the national driver shortage. RIDE should consider awarding the technology/system management contract earlier (e.g., summer 2026), allowing the System Manager platform to be operational before bus operators are onboarded---enabling data-driven route assignments from day one of operations.
\end{description}
