% ===== SECTION 8: TIMELINE =====
\section{\texorpdfstring{\color{MidnightBlue}Timeline}{Timeline}}

\begin{sidebar}{RFI Reference}
Section C: Questions to Field --- Timeline (Question 1, parts a--d)\\[0.3cm]
\textbf{RIDE's Proposed Timeline:}\\
RFP publication: Late spring 2026\\
Responses due: Summer 2026\\
Contract award: November 2026\\
Service start: July 1, 2027 (summer school)
\end{sidebar}

% ----- Question 1a -----
\subsection{Question 1(a): Proposed Timeline Modifications}
\textit{``Is there anything your company or organization would modify within the proposed timeline that would be more conducive to your success?''}

\vspace{0.3cm}

RIDE's proposed timeline is feasible for a technology and system management deployment---Sentry can deploy software, install GPS and camera hardware, integrate with eRIDE, and train operators within an 8-month window. However, the timeline presents substantial risk for bus operators who must simultaneously acquire vehicles, recruit and credential CDL drivers, establish depots, and build route familiarity. Sentry recommends modifications that protect the overall program from the kind of disruption Rhode Island experienced in September 2024, when compressed timelines and inadequate preparation resulted in stranded students and emergency route transfers.

\begin{description}[style=nextline, leftmargin=0.5cm]
    \item[\faIcon{exclamation-triangle} \textbf{8-Month Transition Is Tight:}] The November 2026 award to July 2027 service start provides approximately 8 months for transition. For a technology provider like Sentry, this is feasible---software deployment, API integration with eRIDE, GPS hardware installation, and operator training can be completed within this window. However, for bus operators who must acquire vehicles (6--12 month lead times for new school buses), recruit and credential CDL drivers through the RI ``white card'' process, establish depot facilities, and build route familiarity across unfamiliar zones, 8 months presents significant risk---particularly if a new operator replaces an incumbent. For context, when the General Assembly originally created the Statewide Transportation System in FY~2009, RIDE phased it in over \textit{three full years} (14~LEAs in 2010, 26 in 2011, all LEAs by 2012). The current timeline asks for a full system transition in eight months. The September 2024 Dattco crisis demonstrated what happens when an operator enters service before it is fully prepared.

    \item[\faIcon{calendar-check} \textbf{Recommended Modification:}] Sentry recommends RIDE consider one or more of the following timeline modifications to reduce transition risk:
    \begin{itemize}
        \item Move RFP publication to early spring 2026 (March/April) to give vendors 90+ days for response preparation, allowing more thorough fleet planning and partnership development
        \item Accelerate contract award to September or October 2026 to provide 9--10 months for transition, giving operators additional time for vehicle procurement and driver credentialing
        \item Award the technology/system management contract separately and earlier (e.g., summer 2026) so that the routing, dispatch, GPS tracking, and reporting platform is fully operational before bus operators begin their transition---ensuring that Day One routing data, driver assignments, and vehicle tracking are ready when operators mobilize
        \item Allow a phased service start---use summer school (July--August 2027) as a controlled pilot with reduced route volume, then launch full-year service in September 2027, giving operators two months of live operations to identify and resolve issues before the full student population is transported
    \end{itemize}
\end{description}

% --- Timeline Visualization ---
\vspace{0.3cm}

\begin{center}
\begin{tikzpicture}[
    x=0.75cm, y=-1cm,
    phase/.style={draw=MidnightBlue, fill=MidnightBlue!15, minimum height=0.6cm, anchor=west, font=\scriptsize},
    milestone/.style={diamond, draw=red!80, fill=red!20, minimum size=0.3cm, inner sep=0pt},
    label/.style={font=\scriptsize, anchor=east}
]
    % Month axis
    \foreach \m/\name in {1/Mar,2/Apr,3/May,4/Jun,5/Jul,6/Aug,7/Sep,8/Oct,9/Nov,10/Dec,11/Jan,12/Feb,13/Mar,14/Apr,15/May,16/Jun,17/Jul} {
        \node[font=\tiny, above, rotate=45, anchor=south west] at (\m, -0.2) {\name};
    }
    \node[font=\tiny\bfseries, above] at (5, -0.8) {2026};
    \node[font=\tiny\bfseries, above] at (14, -0.8) {2027};

    % Phases
    \node[phase, minimum width=2.25cm] at (3, 1) {};
    \node[label] at (2.8, 1) {RFP Development};

    \node[phase, minimum width=1.5cm, fill=MidnightBlue!25] at (5, 2) {};
    \node[label] at (2.8, 2) {RFP Published};

    \node[phase, minimum width=2.25cm] at (5, 3) {};
    \node[label] at (2.8, 3) {Vendor Responses};

    \node[phase, minimum width=2.25cm] at (7, 4) {};
    \node[label] at (2.8, 4) {Evaluation};

    \node[phase, minimum width=6cm, fill=MidnightBlue!30] at (9, 5) {};
    \node[label] at (2.8, 5) {\textbf{Transition}};

    \node[phase, minimum width=1.5cm, fill=red!15] at (15, 6) {};
    \node[label] at (2.8, 6) {Summer School};

    \node[phase, minimum width=0.75cm, fill=green!15] at (17, 7) {};
    \node[label] at (2.8, 7) {Full Service};

    % Milestones
    \node[milestone] at (5, 0.3) {};
    \node[font=\tiny, above right] at (5, 0) {RFP Out};

    \node[milestone] at (9, 0.3) {};
    \node[font=\tiny, above right] at (9, 0) {Award};

    \node[milestone] at (15, 0.3) {};
    \node[font=\tiny, above right] at (14.5, 0) {\textbf{Service Start}};

\end{tikzpicture}
\end{center}

% ----- Question 1b -----
\subsection{Question 1(b): Driver Recruitment Timeline}
\textit{``We are aware of the national driver shortage. Does your company or organization feel like the above timeline allows for the necessary time to recruit the number of drivers needed per what is currently outlined in this scope?''}

\vspace{0.3cm}

The proposed timeline presents serious driver recruitment challenges. While Sentry is not a bus operator and does not directly employ CDL school bus drivers, our experience managing workforce pipelines for 150+ transportation providers in the MTA Access-A-Ride program gives us direct insight into the time required to recruit, credential, and deploy transportation workers at scale. The national school bus driver shortage is acute, and Rhode Island's unique credentialing requirements---including the multi-day ``white card'' course at CCRI and DMV-administered driving tests---add weeks to the onboarding timeline that other states do not require.

\begin{description}[style=nextline, leftmargin=0.5cm]
    \item[\faIcon{users} \textbf{Scale of Recruitment Challenge:}] RIDE's STS currently requires approximately 334 drivers plus substitutes across 7 zones. If a new operator wins one or more zones, it may need to recruit 50--150+ CDL-endorsed school bus drivers within 6--8 months. In Rhode Island's current labor market---where drivers are actively recruited by Uber, DoorDash, RIPTA, FedEx, and Amazon, and where the Teamsters have secured a 23\% wage increase over three years for incumbent drivers---this is an extraordinarily challenging timeline. The September 2024 Dattco crisis was fundamentally a driver recruitment failure: Dattco could not field enough RI-credentialed drivers and resorted to deploying CT-based drivers who lacked proper licensing.

    \item[\faIcon{clock} \textbf{CDL Training Timeline:}] A new school bus driver in Rhode Island typically requires 8--12 weeks from initial application to road-ready status: BCI/fingerprint background check (2--3 weeks), CDL classroom and behind-the-wheel training (4--6 weeks), school bus endorsement examination, the RI ``white card'' course at CCRI, DMV driving tests, and route familiarization. Even with the new abbreviated pathway for existing CDL holders, the minimum timeline is 4--6 weeks. This means recruitment must begin no later than March 2027 to have drivers road-ready by July---and that assumes no delays in background checks, DMV scheduling, or training school capacity.

    \item[\faIcon{lightbulb} \textbf{Sentry's Recommendation:}] Sentry recommends RIDE permit pre-award recruitment activities. After short-listing finalists (e.g., September 2026), allow those vendors to begin driver recruitment at their own risk and cost before final contract execution. This ``warm pipeline'' approach ensures that drivers are already progressing through CDL training and RI credentialing when the contract is awarded, reducing the effective mobilization timeline by 2--3 months. Sentry's workforce management platform can support this parallel process by tracking candidate pipelines for multiple operators simultaneously, giving RIDE visibility into each bidder's recruitment readiness as part of the evaluation process.
\end{description}

% ----- Question 1c -----
\subsection{Question 1(c): Number of Routes}
\textit{``How many routes would you likely bid on in an upcoming RFP?''}

\vspace{0.3cm}

Sentry's response to this question depends on the procurement structure RIDE adopts for the eventual RFP. Because Sentry is a technology and system management provider---not a bus operator---our interest is in the system-wide technology layer rather than individual route assignments:

\begin{itemize}
    \item \textbf{As Technology / System Manager:} Sentry would bid to provide technology and system management services across all 334 routes statewide. Our platform's value increases with scale---serving all zones through a single routing, dispatch, tracking, and analytics platform enables cross-zone optimization that is impossible when each operator runs a separate system. This is the model Sentry recommends: a centralized technology contract covering the entire statewide system, upgrading the current System Manager role with modern, real-time infrastructure. In our MTA Access-A-Ride operation, we manage technology across 150+ providers and 10,000+ daily trips---demonstrating that a single platform can serve multiple operators without conflict.
    \item \textbf{As Technology Partner to a Bus Operator:} If RIDE bundles technology into the bus operations contract, Sentry could partner with one or more bus operators to provide the technology platform for their zones. In this scenario, the number of routes would depend on the partner operator's capacity and zone selection. However, Sentry respectfully notes that this model fragments the technology layer across multiple contracts, reduces RIDE's independent oversight, and creates the risk that an operator controls what data RIDE sees about its own performance---the opposite of the transparency this RFI seeks to achieve.
\end{itemize}

% ----- Question 1d -----
\subsection{Question 1(d): Fleet Construct Impact}
\textit{``How does this timeline impact your fleet construct?''}

\vspace{0.3cm}

As a technology provider, Sentry does not procure school buses directly. However, the timeline's impact on fleet construct is a critical risk factor that Sentry's platform is designed to mitigate through fleet tracking, maintenance scheduling, and vehicle lifecycle management. We offer the following observations based on our experience managing fleet logistics for 15,000+ tracked vehicles in the MTA Access-A-Ride system:

\begin{description}[style=nextline, leftmargin=0.5cm]
    \item[\faIcon{bus} \textbf{Vehicle Procurement Lead Times:}] New school bus delivery currently requires 6--12 months from order to delivery, with wheelchair-accessible Type A buses requiring custom modifications that extend lead times further. An 8-month transition period from contract award (November 2026) to service start (July 2027) may not allow sufficient time for operators to order and receive new vehicles. The current fleet includes 30 Type A wheelchair buses and 93 standard Type A buses---vehicle types with the longest procurement lead times. If a new operator replaces an incumbent in any zone, it must either acquire vehicles before the contract is awarded (at its own financial risk) or rely on purchasing or leasing the outgoing operator's fleet.

    \item[\faIcon{video} \textbf{Video System Installation:}] The July 1, 2027 mandate requires all new buses to have live digital video violation detection systems installed from Day One of service. Sentry's camera and GPS hardware can be installed at a rate of 10--15 vehicles per day per installation team, with each vehicle requiring approximately 2--4 hours for camera mounting, wiring, calibration, and system testing. However, installation must be scheduled after vehicle delivery and before route assignment. A compressed timeline increases the risk of buses entering service before camera systems are installed and validated---potentially putting RIDE out of compliance with the new law on the very first day of the contract. Sentry recommends RIDE require operators to provide vehicle delivery schedules 90 days before service start so that technology installation can be pre-scheduled.

    \item[\faIcon{recycle} \textbf{Fleet Transition Strategy:}] Sentry recommends RIDE include a fleet continuity provision in the RFP, requiring outgoing operators to make vehicles available for purchase or lease by incoming operators at fair market value. This approach was used successfully when RIDE transferred 41 routes from Dattco to First Student in 2024---First Student was able to absorb routes in part because existing vehicles remained available. Codifying this provision in the contract eliminates the need for incoming operators to procure an entirely new fleet within 8 months, dramatically reducing transition risk. Sentry's fleet management platform can facilitate this process by maintaining a complete vehicle inventory---including maintenance history, mileage, DOT inspection status, and remaining useful life---so that incoming operators can make informed acquisition decisions immediately upon contract award.
\end{description}
